Long-term Water Quality Change in Rhode Island Lakes and Ponds

### 2019-02-20 manuscript/manuscript.Rmd
added section 1 of 3 to introduction along with associated literature to .bib

1. Water quality in last several decades (Why Water Quality Paragraph(s))
  - Why of interest
    - cyanos stuff
    - CWA stuff
  - Understand long-term trends from local, to regional, to national
  - What variables are of interest and why
    - productivity
    - climate change
    - Nutrient enrichment stuff

>>>>

1. Water quality in last several decades (Why Water Quality Paragraph(s))
  - Why of interest
    - cyanos stuff (toxins, increasing problems)
    - CWA stuff
  - Understand long-term trends from local, to regional, to national
  - What variables are of interest and why
    - productivity
    - climate change
    - Nutrient enrichment stuff

  Aquatic ecosystems have been altered as the result of human activities modifying nutrient cycling on a global scale [@vitousek1997human; @finlay2013human; @filippelli2008global]. Because of their position in the landscape, lakes can function as integrators and sentinels for these anthropogenic effects [@williamson2008lakes; @schindler2009lakes]. Increasing nutrient inputs, particularly of N and P, derived from intensive agriculture and densely populated urban areas have contributed to the eutrophication of many lakes [@carpenter1998nonpoint; @smith2003eutrophication]. Eutrophication has caused the frequency and severity of algal blooms to increase, and these blooms are predicted to worsen under a warming climate [@michalak2013record; @kosten2012warmer; @taranu2015acceleration]. Increasing HABs will cause further declines in water quality, greater risks for human and animal health, and will be economically costly [@brooks2016harmful; @dodds2008eutrophication ;@paerl2009climate]. Management strategies have historically focused on reducing P inputs to lakes, but research also suggests that concurrently reducing N inputs may be more effective in certain situations [@paerl2016takes; @schindler2008eutrophication]. These contrasting studies indicate that spatial differences of HABs exist and that long-term studies are needed to identify trends at local, regional, and national scales.
